---
title: "SDS315 HW7"
author: "Trisha Kundu"
date: "2025-04-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=5, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE, echo=FALSE)
```

## Name: Trisha Kundu

## UT EID: tk24959

## -----------------------


# Problem 1: Armfolding

## A: 
```{r results = FALSE}
library(mosaic)
library(ggplot2)
# read in the dataset
armfold = read.csv("armfold.csv")
# num. of male and female students in dataset
table(armfold$Sex)
# The sample proportion of males who folded their left arm on top (1, male)
male_left_prop <- mean(armfold$LonR_fold[armfold$Sex == "Male"] == 1)
# The sample proportion of females who folded their left arm on top
female_left_prop <- mean(armfold$LonR_fold[armfold$Sex == "Female"] == 1)
```
There are 111 females and 106 males. The proportion of females who cross their left arms over is 0.4234, and 0.472 for the proportion of males who cross their left arms over.


## B. Observed Difference between proportions between males and females
The observed difference in proportions between the two groups (males minus females) is `r (male_left_prop - female_left_prop)`

## C. 95% confidence interval visualized and quantified
```{r}
# 95% conf. int for diff proportions (male-female)
#bootstrapping sample to find difference in mean
armfold_boot <- do(10000) * diffprop(LonR_fold ~ Sex, data = resample(armfold))

#graphing bootstrapped distribution
ggplot(armfold_boot) + geom_histogram(aes(x = diffprop), bins = 30, fill = "lightblue", color = "black") + labs(title = "Distribution of Difference in Proportion", x = "Difference in Proportions (Male - Female)", y = "Frequency")

confint(armfold_boot, level = 0.95)
```
The 95% confidence interval of the difference in proportions (males-females) resulted in the interval of -0.0816 to 1.181. This means that with 95% confidence, the true difference in proportions could range from females crossing left over right more than males (-0.087) to males crossing left over right more than females with a prop. of 1.181. The formula for standard error for the difference in proportions is sqrt((p1 * (1 - p1) / n1) + (p2 * (1 - p2) / n2)). The values we plug in would be p1 = 0.47, p2 = 0.42, n1 = 106, n2 = 111. Plugging them in, the standard error is approximately 0.0674. To find the 95% conf. int., we use 1.96 as the z* value. The formula is the sample mean +/- the (z* value * standard error). Plugging in these values, we get (0.47-0.42) +/- (1.96 * 0.0674) equaling 0.05 +/- 0.1321. The 95% confidence interval from the equation is (-0.082, 0.182). The interval suggests that there may not be a difference in arm preference between genders since the interval includes 0.



## D. Confidence interval in context
If we were to repeat this experiment multiple times, then we would expect that the true difference in the proportion of males and females who cross their arms with their left arm on top would be -0.082 and 0.182, with 95% confidence for the interval.


## E. What does the standard error show?
The standard error calculated above shows the variability we could expect in the difference in proportions between males and females who fold their left arm over right. It measures how much the difference could vary from one sample to another. 

## F. Samping Distribution in context
In this context, sampling distribution is the distribution of difference in proportions we can expect to see after repeated random sampling of males and females who fold their left arm on top from the specific population (Australian university). Sample to sample, the sample proportions for males and females will vary because in each random sample, there will be a different set of people where slight variations in arm folding can take place. On the other hand, the true population proportions stay fixed sample to sample because it is the proportions for all the males and females in the population who fold their left arm on top. They stay fixed since they are the values we are trying to estimate on a larger scale, and we are creating samples from the population. 

## G. The theorem that allows using a normal distribution to approximate the sampling distribution
The Central Limit Theorem justifies using a normal distribution to approximate the sampling distribution of the diff. sample proportions. It tells us that with large enough sample sizes, the distribution of the sample statistics (in this case, difference in proportions) will be approximately normal, even if the shape of the population distribution is not.

## H. Suppose your 95% confidence interval for the difference in proportions was [-0.01, 0.30]. Based on this, what would you say to someone who claims “there’s no sex difference in arm folding”?
I would say that since the interval includes 0, the 95% confidence interval suggests that there may be no difference in the proportion difference of males to females that fold their left arm on top. However, since the positive number (0.30) is a lot higher then -0.01, there may be more males that are likely to fold their left arm on top, so we cannot rule that conclusion out entirely, but we can say that we are 95% confidence that there is no sex difference in arm folding. 

## I. Imagine repeating this experiment many times with different random samples of university students. Would the confidence interval be different across samples? Why? What should be true about the collection of all those intervals?
Yes, the confidence interval would likely be different across samples if we repeated this experiment many times with different random samples of university students. Each random sample drawn from the population of university students will likely contain slightly different proportions. This is because of the random sampling process, so the exact proportions and the  differences in proportions will vary slightly from one sample to another. Since we're sampling randomly, the proportion of males and females who fold their left arm on top will vary depending on which students are included in the sample, which could lead to different estimates of the difference in proportions. The collection of all those intervals would likely contain the true population difference in proportions about 95% of the time.

\newpage

# Problem 2: Get Out the Vote

## Part A. How much more likely are GOTV call recipients to have voted in 1998?

```{r results='hide'}
gotv = read.csv("turnout.csv")
vote98mean = mean(gotv$voted1998[gotv$GOTV_call == 1])
nocall98mean = mean(gotv$voted1998[gotv$GOTV_call == 0])
gotv_no <- mean(gotv$GOTV_call[gotv$voted1998 == 1] == 0)

# 95% confidence interval for the difference in these two proportions
gotv_boot <- do(10000) * diffprop(voted1998 ~ GOTV_call, data = resample(gotv))

# Step 3: Confidence interval
confint(gotv_boot, level = 0.95)
```
The proportion of people who did receive a GOTV call who voted in 1988 is 0.648, and 0.444 for people who voted in 1988 and did not receive a call.
The 95% confidence interval of the difference in these two proportion (received call - did not receive call) is 0.143 to 0.264, which means that the difference in proportions of voting in 1998 for those who received a GOTV call versus those who didn’t is in that interval with 95% confidence.

## Part B. Potential confounders (voted1996, AGE, and MAJORPTY)

```{r}
# variable: voted1966
vote96mean = mean(gotv$voted1996[gotv$GOTV_call == 1])  # treated
nocall96mean = mean(gotv$voted1996[gotv$GOTV_call == 0])  # control
vote1966boot = do(10000) * diffmean(voted1996 ~ GOTV_call, data = resample(gotv))
confint(vote1966boot, level=0.95)
```
## Voting in 1996
Among individuals who received a GOTV call, 71.2% had voted in the 1996 election, compared to only 53.1% of those who did not receive a call. The difference in proportions, with a 95% confidence interval ranges from 0.124 to 0.237, indicating that this difference is statistically significant since 0 is not in the interval.
We see that people who received calls were already more likely to vote since they voted in 1996, making voting in 1996 a confounding variable

```{r}
# variable: age
matchage = mean(gotv$AGE[gotv$GOTV_call == 1])
matchnoage = mean(gotv$AGE[gotv$GOTV_call == 0])
age_boot = do(10000) * diffmean(AGE ~ GOTV_call, data = resample(gotv))
confint(age_boot, level = 0.95)

ggplot(gotv) + geom_boxplot(aes(x = factor(GOTV_call), y = AGE, fill = factor(GOTV_call))) + facet_wrap(~voted1998, labeller = labeller(voted1998 = c("0" = "Did Not Vote in 1998", "1" = "Voted in 1998"))) + labs(x = "GOTV Call (0 = No, 1 = Yes)", y = "Age", fill = "GOTV Call", title = "Age by GOTV Call and 1998 Voting")
```
## Age
The mean age of individuals who received a GOTV call is 58.3 years old, and 49.42 years old for people who did not receive a call.
The bootstrapped 95% confidence level for the difference in means for the age of voters with vs. without a GOTV call is between 6.429 to 11.343. Since 0 is not in the confidence interval, the difference in means is significant, meaning age does factor into the data and is a confounding variable affecting the voting outcome.


```{r}

# variable: MAJORPTY
mjr_gotv=mean(gotv$MAJORPTY[gotv$GOTV_call == 1])
nomjr_gotv=mean(gotv$MAJORPTY[gotv$GOTV_call == 0])
major_boot = do(10000) * diffmean(MAJORPTY ~ GOTV_call, data = resample(gotv))
confint(major_boot, level=0.95)

```
## Voting for a major party
Among those who received a GOTV call, 80.2% were registered with a major party, compared to 74.5% of those who didn’t. The 95% confidence interval for this difference is 0.00531 to 0.1059, meaning there is a significant association. Since major party members were both more likely to get a call and more likely to vote, MAJORPTY is a confounder that affects the estimate of the true effect of the GOTV call.


# Part C.
```{r results='hide'}
library(tidyverse)
library(mosaic)
library(MatchIt)

voting_match = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = gotv, ratio=5)

gotv_matched = match.data(voting_match)

#show it is balanced with 95% conf int.
# Voted1996
confound96 = mean(voted1996 ~ GOTV_call, data = gotv_matched)
match1996_boot <- do(10000) * diffprop(voted1996 ~ GOTV_call, data = resample(gotv_matched))
confint(match1996_boot, level = 0.95)

# AGE 
confoundage = mean(AGE ~ GOTV_call, data = gotv_matched)
matchage_boot <- do(10000) * diffmean(AGE ~ GOTV_call, data = resample(gotv_matched))
confint(matchage_boot, level = 0.95)

# 3. MAJORPTY 
confoundmajor = mean(MAJORPTY ~ GOTV_call, data = gotv_matched)
matchmajor_boot <- do(10000) * diffprop(MAJORPTY ~ GOTV_call, data = resample(gotv_matched))
confint(matchmajor_boot, level = 0.95)

# more analysis from part A
confound98 = mean(voted1998 ~ GOTV_call, data =gotv_matched)
voted_boot <- do(10000) * diffprop(voted1998 ~ GOTV_call, data = resample(gotv_matched))
confint(voted_boot, level = 0.95)
```

```{r}

ggplot(voted_boot) + geom_histogram(aes(x = diffprop), bins = 30, fill = "lightblue", color = "black") + labs(title = "Bootstrapped Distribution of Difference in Proportion", x = "Difference in Proportions (GOTV Call vs. No GOTV Call)", y = "Frequency")
```

By matching voted1996, AGE, and MAJORPTY using a ratio=5 to balance the data set. After matching, all three confounders were balanced and no longer confounders for the data set. We know this because in the matched data, when computing 95% confidence intervals between the confounders vs. getting a GOTV call, zero is included in the data set, showing there is no significance between the 2 variables. The confidence interval for the difference in proportions of voting in 1996 and getting a GOTV call is -0.0629 to 0.0618, with 95% confidence. For age vs. GOTV call, it is -2.654 to 2.727, with 95% confidence. For major party vs. getting a GOTV call, the interval for difference in proportions is -0.0529 to 0.047, with 95% confidence. Since 0 is included in all 3 confidence interval comparing confounders, we can say that the matched data set is balanced.

The proportions of those who received a GOTV call who voted in 1988 is 56.92%. The sample proportion of those not receiving a GOTV call who voted in 1998 is 64.78%. The large-sample 95% confidence interval for the difference in these 2 proportions is between 0.01304 to 0.14511, with 95% confidence.

Using the data analysis, the GOTV call appears to have a positive effect on the likelihood of voting in the 1998 election. The confidence interval can suggest that receiving a GOTV call increases the likelihood of someone voting. This is because the confidence interval does not include zero, meaning there is statistically significance between receiving a GOTV call and voting in 1998.




